---
title: "DS6372_Project2"
author: "Samuel Onalaja"
date: "11/11/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library

```{r}
library(tidyverse)
library(corrplot)
library(mlbench)
library(caret)
library(skimr)
library(mice)
library(purrr)
library(ggplot2)
library(ggthemes)
library(cowplot)
library(class)
library(e1071)
library(grid)
library(tidyr)
library(stringr)
library(naniar)
library(car)

```

Read in datafile
```{r}

library(readr)
bank_additional_full <- read_delim("data/bank-additional-full.csv", ";", escape_double = FALSE, trim_ws = TRUE)
head(bank_additional_full)
View(bank_additional_full)
```

Data Summary statistics

Notice all categorical variables are in charactor type so we have to convert all to factor.
```{r}
summary(bank_additional_full)
dim(bank_additional_full)
names(bank_additional_full)
str(bank_additional_full)

```


Convert all character type variable to factor
```{r}

for (i in seq_along(bank_additional_full)){
    if(is.character(bank_additional_full[[i]])){
        bank_additional_full[[i]]=as.factor(bank_additional_full[[i]])   
    }
}

str(bank_additional_full)
```

Addressing missing value

The data set is completely observed, No missing value
```{r}

vis_miss(bank_additional_full) #This funtion is from package naniar and very effective for visualizing missing values
md.pattern(bank_additional_full) #This function also address the pattern of a missing value, its from the The MICE package (stands for Multiple Imputation by Chained Equations)

n_cols = dim(bank_additional_full)[2]
for (i in 1:n_cols)
{
  print(c(colnames(bank_additional_full[i]), sum(is.na(bank_additional_full[i]))))
}

```



EDA

Notice some errors while trying to get a correlation so it was discovered that the record "No and Yes" in the "y" variable needs to be converted to "0" and "1"

Notice how the respose variable (y) is skewed towards "0" which is no at over 89%

```{r}


bank <- bank_additional_full %>%
  mutate(y = ifelse(y=="yes", 1, 0))

bank$y <- as.factor(bank$y)

table(bank$y)
prop.table(table(bank$y))

dim(bank)
str(bank)

#Observe data distribution of the response variable

table(bank$y)
prop.table(table(bank$y))


## Check of the range of some of these continuous variables with histograms
 
#Distribution of duration

ggplot(data=bank,aes(x=duration)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$duration, na.rm = TRUE), sd = sd(bank$duration, na.rm = TRUE))) +
  labs(x = "duration", y = "Count", title = "Duration distribution") +   theme_economist()

summary(bank$duration)

#Campaign distribution
#Campaign is rightly skewed and contains some few outliers

ggplot(data=bank,aes(x=campaign)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$campaign, na.rm = TRUE), sd = sd(bank$campaign, na.rm = TRUE))) +
  labs(x = "campaign", y = "Count", title = "Campaign distribution") +   theme_economist()

#checking out the distribution by the response variable
#Distribution seems to be normally distributed
ggplot(bank) + geom_histogram(aes(x = age), binwidth = 0.1, col = "white") + facet_grid(y~., scales = "free") + scale_x_log10() +theme_bw()

summary(bank$campaign)

bank$lgcampaign = log(bank$campaign)

#Pdays distribution

ggplot(data=bank,aes(x=bank$pdays)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$campaign, na.rm = TRUE), sd = sd(bank$pdays, na.rm = TRUE))) +
  labs(x = "pdays", y = "Count", title = "Pdays distribution") +   theme_economist()


#Previous distribution

ggplot(data=bank,aes(x=bank$previous)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$previous, na.rm = TRUE), sd = sd(bank$previous, na.rm = TRUE))) +
  labs(x = "previous", y = "Count", title = "Previous distribution") +   theme_economist()


#Emp.var.rate distribution

ggplot(data=bank,aes(x=bank$emp.var.rate)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$emp.var.rate, na.rm = TRUE), sd = sd(bank$emp.var.rate, na.rm = TRUE))) +
  labs(x = "emp.var.rate", y = "Count", title = "Emp.var.rate distribution") +   theme_economist()


#Cons.price.idx distribution

ggplot(data=bank,aes(x=bank$cons.price.idx)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$emp.var.rate, na.rm = TRUE), sd = sd(bank$cons.price.idx, na.rm = TRUE))) +
  labs(x = "cons.price.idx", y = "Count", title = "Cons.price.idx distribution") +   theme_economist()


#Cons.conf.idx distribution

ggplot(data=bank,aes(x=bank$cons.conf.idx)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$cons.conf.idx, na.rm = TRUE), sd = sd(bank$cons.conf.idx, na.rm = TRUE))) +
  labs(x = "cons.conf.idx", y = "Count", title = "Cons.conf.idx distribution") +   theme_economist()


#Euribor3m distribution

ggplot(data=bank,aes(x=bank$euribor3m)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$euribor3m, na.rm = TRUE), sd = sd(bank$euribor3m, na.rm = TRUE))) +
  labs(x = "euribor3m", y = "Count", title = "Euribor3m distribution") +   theme_economist()


#Nr.employed distribution

ggplot(data=bank,aes(x=bank$nr.employed)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$nr.employed, na.rm = TRUE), sd = sd(bank$nr.employed, na.rm = TRUE))) +
  labs(x = "nr.employed", y = "Count", title = "Nr.employed distribution") +   theme_economist()





#checking for correlation Numerical variable vs response

Attr <- "y"
# Name explanatory variable
ExplVar<- bank%>% keep(is.numeric) %>% colnames
# Create function
PlotFunc <- function(df, explanatory, response) {
  ggplot(data = df) + geom_density(aes_string(x = explanatory), alpha = 0.5) + xlab(explanatory) + ylab("subscribed")
}
  # Density plot
PlotFunc(bank, explanatory =  "age", response = "y")
#  Create plot list for plot_grid function to reference
PlotList <- lapply(ExplVar, function(x) PlotFunc(bank, x, y))
#  Grid of all categorical variables plotted against Attrition
plot_grid(plotlist = PlotList)



```


This table shows the correlation between the numerical variables

 - nr.employed and emp.var.rate are 91% correlated. 
 - nr.employed and euribor3m are 95% correlated.
 - emp.var.rate and euribor3m are 97% correlated.
 - cons.price.idx and emp.var.rate are 78% correlated.
 - cons.price.idx and euribor3m are 69% correlated.
 - ons.price.idx and nr.employed are 52% correlated.
 
```{r}


corrdfTraintable <- bank %>% keep(is.numeric) %>% na.omit %>% cor %>% view
bank %>% keep(is.numeric) %>% na.omit %>% cor %>% corrplot("upper", addCoef.col = "black", number.digits = 2, number.cex = 0.5, method="shade", order="hclust", tl.srt=45, tl.cex = 0.8)
view(corrdfTraintable)

pairs

plot_correlation(bank, type ='continous', 'Review.Date')
```



Using the package SmartEda for data exploration

- check out the distributiion of numerical variable and categorical variable.

```{r}
library("DataExplorer")
library('SmartEDA')

ExpData(data = bank, type = 1)

BankPlot <- ExpNumViz(bank, target = NULL, nlim = 10, Page = c(3,3), sample = 9)

BankPlotn <- ExpNumViz(bank, target = "y", nlim = 10, Page = c(3,3), sample = 9)

Bankplot2 <- ExpCatViz(bank,target= "y",col ="slateblue4",clim=10,margin=2,Page = c(3,3),sample=9)


pairs(bank[,c(1,11:14,16:20)], col = bank$y)

cor(bank[,c(1,11:14,16:20)], col = bank$y)
```

Missing values are disguided as unknwown values in the data set and we devised a code to show them all

 default           8597
 education         1731
 housing            990
 loan               990
 job                330
 marital             80
 
 For Marital, housing and job it is save to remove the unknow values as the so little they won't have an effect on the rest of the disrtibution
 
 - Looking at default is has no information as it is highly skewed towards "no" as only 3 counts comes up as "yes" so we are removing "default" as well.
 
 
 
 
```{r}
bank %>% 
  summarise_all(list(~sum(. == "unknown"))) %>% 
  gather(key = "variable", value = "nr_unknown") %>% 
  arrange(-nr_unknown)

summary(bank$job)
summary(bank$default)
summary(bank$education)
summary(bank$loan)
summary(bank$marital)
summary(bank$housing)

bank <- subset(bank, job!="unknown")
bank <- subset(bank, marital!="unknown")
bank <- subset(bank, housing!="unknown")
bank <- subset(bank, loan!="unknown")
bank <- subset(bank, default!="unknown")
bank <- subset(bank, education!="unknown")




###cov.trob(bankbank[,c(1,11:14,16:20)])$cov
```


```{r}
BankBoxPlot <- function(columnBy, columnToShow, columnByDesc, columnToShowDesc)
{
ggplot(bank, aes(x=bank[,columnBy], y=bank[,columnToShow])) + geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", shape=5, size=4) +
  ggtitle(paste(columnByDesc, "by",columnToShowDesc)) + xlab("Outcome") + ylab(columnToShowDesc) +
  theme_economist()
}
BankBoxPlot("y", "duration", "Outcome", "duration - Duration of Last Contact (secs)")
BankBoxPlot("y", "campaign", "Outcome", "campaign - Number of Contacts During this Campaign")
BankBoxPlot("y", "pdays", "Outcome", "pdays - Number of Days Since Previous Contact")
BankBoxPlot("y", "previous", "Outcome", "previous - Previous Number of Contacts")
NiceBoxPlot("y", "emp.var.rate", "Outcome", "emp.var.rate - Employment Variation Rate (quaterly indicator)")
NiceBoxPlot("y", "cons.price.idx", "Outcome", "cons.price.idx - Consumer Price Index (montly indicator)")
NiceBoxPlot("y", "cons.conf.idx", "Outcome", "cons.conf.idx - Consumer Confidence Index(monthly indicator)")
NiceBoxPlot("y", "euribor3m", "Outcome", "euribor3m - eurobor 3 month rate")
NiceBoxPlot("y", "nr.employed", "Outcome", "nr.employed - number of employees (quarterly indicator)")
```

```{r}

library("dplyr")
ShowPieChart <- function(columnBy, columnToShow, titleName)
{
  df <- dplyr::group_by(bank, .dots = c(columnBy, columnToShow)) %>%
    dplyr::summarize(counts = n()) %>%
    dplyr::mutate(perc = (counts / sum(counts)) * 100) %>%
    dplyr::arrange_(.dots=c(columnBy, columnToShow))
 
  
  # preparing the plot
  ggplot2::ggplot(df, aes('', counts)) +
    geom_col(
      position = 'fill',
      color = 'black',
      width = 1,
      aes(fill = y)
    ) +
    ggtitle(titleName) +
    facet_wrap(paste("~",columnBy), labeller = "label_both") +
    geom_label(
      aes(label = paste0(round(perc), "%"), group = "y"),
      position = position_fill(vjust = 0.5),
      color = 'black',
      size = 5,
      show.legend = FALSE
    ) + scale_fill_discrete(name = "Outcome:") +
    coord_polar(theta = "y")
}
ShowPieChart("job", "y", "Outcome by Job")
ShowPieChart("marital", "y", "Outcome by Marital Status")
ShowPieChart("education", "y", "Outcome by Education")
ShowPieChart("housing", "y", "Outcome by Housing")
ShowPieChart("default", "y", "Outcome by Credit In Default")
ShowPieChart("loan", "y", "Outcome by loan status")
ShowPieChart("contact", "y", "Outcome by Contact")
ShowPieChart("poutcome", "y", "Outcome by poutcome")
 
```



```{r}


## Check of the range of some of these continuous variables with histograms
 

ggplot(data=bank,aes(x=duration)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$duration, na.rm = TRUE), sd = sd(bank$duration, na.rm = TRUE))) +
  labs(x = "duration", y = "Count", title = "Distribution of duration") +   theme_economist()
ggplot(data=bank,aes(x=bank$campaign)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$campaign, na.rm = TRUE), sd = sd(bank$campaign, na.rm = TRUE))) +
  labs(x = "campaign", y = "Count", title = "Distribution of campaign") +   theme_economist()
ggplot(data=bank,aes(x=bank$pdays)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$campaign, na.rm = TRUE), sd = sd(bank$campaign, na.rm = TRUE))) +
  labs(x = "pdays", y = "Count", title = "Distribution of pdays") +   theme_economist()
ggplot(data=bank,aes(x=bank$previous)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$previous, na.rm = TRUE), sd = sd(bank$campaign, na.rm = TRUE))) +
  labs(x = "previous", y = "Count", title = "Distribution of previous") +   theme_economist()
ggplot(data=bank,aes(x=bank$emp.var.rate)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$emp.var.rate, na.rm = TRUE), sd = sd(bank$emp.var.rate, na.rm = TRUE))) +
  labs(x = "emp.var.rate", y = "Count", title = "Distribution of emp.var.rate") +   theme_economist()
ggplot(data=bank,aes(x=bank$cons.price.idx)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$emp.var.rate, na.rm = TRUE), sd = sd(bank$cons.price.idx, na.rm = TRUE))) +
  labs(x = "cons.price.idx", y = "Count", title = "Distribution of cons.price.idx") +   theme_economist()
ggplot(data=bank,aes(x=bank$cons.conf.idx)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$cons.conf.idx, na.rm = TRUE), sd = sd(bank$cons.conf.idx, na.rm = TRUE))) +
  labs(x = "cons.conf.idx", y = "Count", title = "Distribution of cons.conf.idx") +   theme_economist()
ggplot(data=bank,aes(x=bank$euribor3m)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$euribor3m, na.rm = TRUE), sd = sd(bank$euribor3m, na.rm = TRUE))) +
  labs(x = "euribor3m", y = "Count", title = "Distribution of euribor3m") +   theme_economist()
ggplot(data=bank,aes(x=bank$nr.employed)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(bank$nr.employed, na.rm = TRUE), sd = sd(bank$nr.employed, na.rm = TRUE))) +
  labs(x = "nr.employed", y = "Count", title = "Distribution of nr.employed") +   theme_economist()
```

